---
title: "Geospatial Package Notes"
author: "Fisher Ankney"
date: "3/8/2018"
output: html_document
---

Important Packages for Geospatial Data - 

ggmap - basic map download from several sources
ggplot2 - useful with ggmap, basic maps 

sp - classes and methods for spatial data
tmap - really good mapping package

raster - good for raster data

RColorBrewer - colors
Viridislite - colors 

rgdal - bindings for geospatial data 
tigris - download and use census shapefiles


```{r}
library('ggmap')
corvallis <- c(lon = -123.2620, lat = 44.5646)

# Get map at zoom level 5: map_5
map_5 <- get_map(corvallis, zoom = 5, scale = 1)

# Plot map at zoom level 5
ggmap(map_5)

# Get map at zoom level 13: corvallis_map
corvallis_map <- get_map(corvallis, zoom = 13, scale = 1)

# Plot map at zoom level 13
ggmap(corvallis_map)
```

you can add layers of data to ggmap, call +geom_point() to overlay data points onto a map. 

```{r}
# Look at head() of sales
head(sales)

# Swap out call to ggplot() with call to ggmap()
# ggmap(corvallis_map) +
#  geom_point(aes(lon, lat), data = sales)
```


```
# Map color to price / finished_squarefeet
ggmap(corvallis_map) +
  geom_point(aes(lon, lat, color = price / finished_squarefeet), data = sales)
```

source and type of map default is maptype = "terrain", source = "google"

extent controls how much of the plotting area the map takes up, plot range also important in ggmap()

types of data - 
- point data <br>
- line data <br> 
- polygon data <br> 
- raster data <br> 

SP package - 
spatial package objects for point, line, polygon data 
sp objects important to master

Practice accessing slots, an object oriented type of 



```
# Create logical vector: is_nz
is_nz <- countries_spdf$name == "New Zealand"

# Subset countries_spdf using is_nz: nz
nz <- countries_spdf[[is_nz, ]]

# Plot nz
plot(nz)
```


tmap package
```{r}
library(tmap)
```

```{r}
tm_shape(Europe) + 
  tm_borders() + 
  tm_fill(col = "part") + 
  tm_compass() + 
  tmap_style("cobalt")
```

so ggplot2 and tmap are similiar, but ggplot2 is designed for dataframe data, tmap expect data to be stored in a spatial object. each layer contains a mapping variable. 

tmap provides no functions equivilant to the scale functions, all tweaks are in the layer call, tm_shape defines data for any subsequent layer, multiple calls possible

```
library(sp)
library(tmap)

# Add style argument to the tm_fill() call
tm_shape(countries_spdf) +
  tm_fill(col = "population", style = "quantile") +
  # Add a tm_borders() layer 
  tm_borders(col = "burlywood4")

# New plot, with tm_bubbles() instead of tm_fill()
tm_shape(countries_spdf) + 
tm_bubbles(size = "population") + 
tm_borders(col = "burlywood4")

```


```
library(sp)
library(tmap)

# Switch to a Hobo–Dyer projection
tm_shape(countries_spdf, projection = "hd") +
  tm_grid(n.x = 11, n.y = 11) +
  tm_fill(col = "population", style = "quantile")  +
  tm_borders(col = "burlywood4") 

# Switch to a Robinson projection
tm_shape(countries_spdf, projection = "robin") +
  tm_grid(n.x = 11, n.y = 11) +
  tm_fill(col = "population", style = "quantile")  +
  tm_borders(col = "burlywood4") 

# Add tm_style_classic() to your plot
tm_shape(countries_spdf, projection = "robin") + 
tm_grid(n.x = 11, n.y = 11) + 
tm_fill(col = "population", style = "quantile") + 
tm_borders(col = "burlywood4") + 
tm_style_classic()
```

fun idea! plot the countries ive been to using this old world style! 

```
library(sp)
library(tmap)

# Plot from last exercise
tm_shape(countries_spdf) +
  tm_grid(n.x = 11, n.y = 11, projection = "longlat") +
  tm_fill(col = "population", style = "quantile")  +
  tm_borders(col = "burlywood4")

# Save a static version "population.png"
save_tmap(filename = "populaton.png")

# Save an interactive version "population.html
save_tmap(filename = "population.html")
```

### The raster package 

raster provides print methods for sp objects, cool

```
summary(pop)

str(pop)

# Call plot() on pop
plot(pop)

# Call str() on values(pop)
val_pop <- values(pop)
str(val_pop)


# Call head() on values(pop)
head(val_pop)
```

```
# Print pop_by_age
pop_by_age

# Subset out the under_1 layer using [[
pop_by_age[["under_1"]]

# Plot the under_1 layer
plot(pop_by_age[["under_1"]])
```

RColorBrewer good for maps

colors 

```
library(RColorBrewer)
# 9 steps on the RColorBrewer "BuPu" palette: blups
blups <- brewer.pal(9, "BuPu")

# Add scale_fill_gradientn() with the blups palette
ggplot(preds) +
  geom_tile(aes(lon, lat, fill = predicted_price), alpha = 0.8) +
  scale_fill_gradientn(colors = blups)

library(viridisLite)
# viridisLite viridis palette with 9 steps: vir
vir <- viridis(9)

# Add scale_fill_gradientn() with the vir palette
ggplot(preds) +
  geom_tile(aes(lon, lat, fill = predicted_price), alpha = 0.8) +
  scale_fill_gradientn(colors = vir)
  
# mag: a viridisLite magma palette with 9 steps
mag <- magma(9)

# Add scale_fill_gradientn() with the mag palette
ggplot(preds) +
  geom_tile(aes(lon, lat, fill = predicted_price), alpha = 0.8) +
  scale_fill_gradientn(colors = mag)  
```

library('classInt')


```
mag <- viridisLite::magma(7)

library(classInt)

# Create 5 "pretty" breaks with classIntervals()
classIntervals(values(prop_by_age$age_18_24), n = 5, style = "pretty")

# Create 5 "quantile" breaks with classIntervals()
classIntervals(values(prop_by_age$age_18_24), n = 5, style = "quantile")



# Use 5 "quantile" breaks in tm_raster()
tm_shape(prop_by_age) +
  tm_raster("age_18_24", palette = mag, n = 5, style = "quantile") +
  tm_legend(position = c("right", "bottom"))

# Create histogram of proportions
hist(values(prop_by_age)[, "age_18_24"])

# Use fixed breaks in tm_raster()
tm_shape(prop_by_age) +
  tm_raster("age_18_24", palette = mag,
    style = "fixed", breaks = c(0.025, 0.05, 0.1, 0.2, 0.25, 0.3, 1))

# Save your plot to "prop_18-24.html"
save_tmap(filename = "prop_18-24.html")
```

```
# Print migration
migration

# Diverging "RdGy" palette
red_gray <- brewer.pal(n = 7, "RdGy")

# Use red_gray as the palette 
tm_shape(migration) +
  tm_raster(palette = red_gray) +
  tm_legend(outside = TRUE, outside.position = c("bottom"))

# Add fixed breaks 
tm_shape(migration) +
  tm_raster(palette = red_gray,style = "fixed", breaks = c(-5e6, -5e3, -5e2, -5e1, 5e1, 5e2, 5e3, 5e6)) +
  tm_legend(outside = TRUE, outside.position = c("bottom"))
  ```
  
  
```
library(raster)

# Plot land_cover
tm_shape(land_cover) +
  tm_raster() 

# Palette like the ggplot2 default
hcl_cols <- hcl(h = seq(15, 375, length = 9), 
                c = 100, l = 65)[-9]

# Use hcl_cols as the palette
tm_shape(land_cover) +
  tm_raster(palette = hcl_cols) 

# Examine levels of land_cover
levels(land_cover)

# A set of intuitive colors
intuitive_cols <- c(
  "darkgreen",
  "darkolivegreen4",
  "goldenrod2",
  "seagreen",
  "wheat",
  "slategrey",
  "white",
  "lightskyblue1"
)

# Use intuitive_cols as palette
tm_shape(land_cover) +
  tm_raster(palette = intuitive_cols) +
  tm_legend(position = c("left", "bottom"))
  ```
  
  
  reading shape files into R 
  
library(sp)
library(rgdal)

readOGR() from the rgdal package
two arguements, dsn and layer


```
library(sp)
library(rgdal)

# Use dir() to find directory name
dir()

# Call dir() with directory name
dir("nynta_16c")

# Read in shapefile with readOGR(): neighborhoods
neighborhoods <- readOGR("nynta_16c", "nynta")

# summary() of neighborhoods
summary(neighborhoods)
```
taken from here - https://www1.nyc.gov/site/planning/data-maps/open-data/dwn-nynta.page


reading in a raster from here - http://sedac.ciesin.columbia.edu/data/collection/usgrid

```
library(raster) 

# Call dir()
dir()

# Call dir() on the directory
dir("nyc_grid_data")

# Use raster() with file path: income_grid
income_grid <- raster("nyc_grid_data/m5602ahhi00.tif")

# Call summary() on income_grid
summary(income_grid)

# Call plot() on income_grid
plot(income_grid)
```


data comes from packages nicely 
rnaturalearth package is good 
tigris package provides a way to easily download and import shapefiles based on US Census geographies 
```
library(sp)
library(tigris)

# Call tracts(): nyc_tracts
nyc_tracts <- tracts(state = "NY", county = "New York", cb = TRUE)

# Call summary() on nyc_tracts
summary(nyc_tracts)

# Plot nyc_tracts
plot(nyc_tracts)
```


conforming coordinate systems using sptransform 

```
library(sp)

# proj4string() on nyc_tracts and neighborhoods
proj4string(neighborhoods)
proj4string(nyc_tracts)

# coordinates() on nyc_tracts and neighborhoods
coordinates(nyc_tracts)
coordinates(neighborhoods)


# plot() neighborhoods and nyc_tracts
plot(neighborhoods)
plot(nyc_tracts)
```


```
library(sp)
library(raster)

# Use spTransform on neighborhoods: neighborhoods
neighborhoods <- spTransform(neighborhoods, CRS = proj4string(nyc_tracts))


# head() on coordinates() of neighborhoods
head(coordinates(neighborhoods))

# Plot neighborhoods, nyc_tracts and water
plot(neighborhoods)
plot(nyc_tracts, add = T, col = "red")
plot(water, add = T, col = "blue")
```


```
library(sp)

# Use str() on nyc_income and nyc_tracts@data
str(nyc_income)
str(nyc_tracts@data)


# Highlight tract 002201 in nyc_tracts
plot(nyc_tracts)
plot(nyc_tracts[nyc_tracts$TRACTCE == "002201", ], 
     col = "red", add = TRUE)

# Set nyc_tracts@data to nyc_income
nyc_tracts@data = nyc_income

# Highlight tract 002201 in nyc_tracts
plot(nyc_tracts)
plot(nyc_tracts[nyc_tracts$tract == "002201", ], col = "red", add = TRUE)
```

sometimes you duplicate entries on accident, check for this

```
# Check for duplicates in nyc_income
any(duplicated(nyc_income$tract))

# Check for duplicates in nyc_tracts
any(duplicated(nyc_tracts$TRACTCE))

# Check nyc_tracts in nyc_income
all(nyc_tracts$TRACTCE %in% nyc_income$tract)

# Check nyc_income in nyc_tracts
all(nyc_income$tract %in% nyc_tracts$TRACTCE)
```

```
library(sp)
library(tmap)

# Merge nyc_tracts and nyc_income: nyc_tracts_merge
nyc_tracts_merge <- merge(nyc_tracts, nyc_income, by.x = "TRACTCE", by.y = "tract")

# Call summary() on nyc_tracts_merge
summary(nyc_tracts_merge)

# Choropleth with col mapped to estimate
tm_shape(nyc_tracts_merge) + 
tm_fill(col = "estimate")
```

```
library(tmap)

tm_shape(nyc_tracts_merge) +
  tm_fill(col = "estimate") +
  # Add a water layer, tm_fill() with col = "grey90"
  tm_shape(water) +
  tm_fill(col = "grey90") +
  # Add a neighborhood layer, tm_borders()
  tm_shape(neighborhoods) +
  tm_borders() 
```


```
library(tmap)

# gsub() to replace " " with "\n"
manhat_hoods$name <- gsub(" ", "\n", manhat_hoods$NTAName)

# gsub() to replace "-" with "/\n"
manhat_hoods$name <- gsub("-", "/\n", manhat_hoods$name)

# Edit to map text to name, set size to 0.5
tm_shape(nyc_tracts_merge) +
    tm_fill(col = "estimate") +
  tm_shape(water) +
    tm_fill(col = "grey90") +
  tm_shape(manhat_hoods) +
    tm_borders() +
    tm_text(text = "name", size = 0.5)
```

```
library(tmap)

tm_shape(nyc_tracts_merge) +
  # Add title and change palette
  tm_fill(col = "estimate", 
          title = "Median Income",
          palette = "Greens") +
  # Add tm_borders()
  tm_borders(col = "grey60", lwd = 0.5) +
  tm_shape(water) +
  tm_fill(col = "grey90") +
  tm_shape(manhat_hoods) +
  # Change col and lwd of neighborhood boundaries
  tm_borders(col = "grey40", lwd = 2) +
  tm_text(text = "name", size = 0.5) +
  # Add tm_credits()
  tm_credits("Source: ACS 2014 5-year Estimates, \n accessed via acs package", 
             position = c("right", "bottom"))

# Save map as "nyc_income_map.png"
save_tmap(filename = "nyc_income_map.png", width = 4, height = 7)
```



