---
title:
author: "Fisher Ankney"
date: "3/11/2018"
output: html_document
---

### Greece

insert image of greek flag here

One of my favorite countries, visited twice, plan on going again whenever I get the chance. Perfect opportuntiy to test out new found spatial analysis skills in R on one of my favorite destinations. Understanding greece using technologies in R. Who said you need GIS? 

```{r, message=F}
library('tidyverse')    
library('ggmap')
library('sp')
library('tmap')
library('rgdal')
library('tigris')
library('RColorBrewer')
library('classInt")
```

<br> 

If you forgot where Greece is
```{r, message=F}
greece <- c(lon = 21.8243, lat = 39.0742)

greece_far <- get_map(greece, zoom = 5, scale = 1, maptype="terrain")

ggmap(greece_far)
```

<br> 

Greece is broken up into 11 adm

```{r}
#dir("input/periphereies")


grc_shape <- readOGR("input/periphereies", "periphereies")

summary(grc_shape)

str(grc_shape, max.level = 2)

tm_shape(grc_shape) + 
  tm_borders(lwd = 1, col = "#ffffff") + 
  tm_fill(col = "#0D5EAF") 
```

http://geodata.gov.gr/en/dataset?tags=administrative-units

<br> 

but where to stay? I heard it hard to get a room in athens and it's really expensive, is this true? 
```{r}
athens_airbnb <- read_csv("input/listings (1).csv")
athens_coord <- c(lon = 23.7275, lat = 37.9838)

classIntervals(athens_airbnb$price, n = 5, style = "quantile")
breaks <- c(9,25,35,46,69,5000)
labels <- c("$9 - $25", "$25 - $35", "$35 - $46", "$46 - $69", "$69+")
athens_airbnb$price <- cut(athens_airbnb$price, breaks = breaks, labels = labels)
athens_airbnb$price

athens <- get_map(athens_coord, zoom = 13, scale = 1, maptype="terrain")

ggmap(athens) + 
  geom_point(data = athens_airbnb, 
             aes(longitude, latitude, col = price), 
             alpha = 0.3) 
```
http://insideairbnb.com/get-the-data.html



Wow, can't wait to visit greece. also can't wait to learn more spatial analysis skills with R, really awesome! 















Perhaps The next time I visit I'll go to the biggest island of all, Crete. 
```{r}
crete <- c(lon = 24.8093, lat = 35.2401)

crete_far <- get_map(crete, zoom = 5, scale = 1, maptype="terrain")

ggmap(crete_far)

# Get map at zoom level 13: corvallis_map
crete_map <- get_map(crete, zoom = 8, scale = 1)

# Plot map at zoom level 13
ggmap(crete_map)

```


many archeology sites on crete
```{r}
# arceology in Crete (DATA RETRIEVED)
```
<br> 

and I better hurry because they're threatened by earthquakes
```{r}
# 
earthquakes and archeology sites 
```
https://www.apdkritis.gov.gr/el/dataset/%CE%B1%CF%81%CF%87%CE%B1%CE%B9%CE%BF%CE%B9-%CE%BD%CE%B1%CE%BF%CE%B9
https://www.apdkritis.gov.gr/el/dataset/%CE%B1%CF%81%CF%87%CE%B1%CE%B9%CE%BF%CE%BB%CE%BF%CE%B3%CE%B9%CE%BA%CE%BF%CE%B9-%CF%87%CF%89%CF%81%CE%BF%CE%B9
https://www.apdkritis.gov.gr/el/dataset/%CE%B1%CE%BD%CE%B1%CE%BA%CF%84%CE%BF%CF%81%CE%B1
earthquakes - https://www.apdkritis.gov.gr/el/dataset/%CF%83%CE%B5%CE%B9%CF%83%CE%BC%CE%BF%CE%B9-%CF%83%CF%84%CE%B7%CE%BD-%CE%BA%CF%81%CE%B7%CF%84%CE%B7

<br> 

but where to stay? I heard it hard to get a room in athens and it's really expensive, is this true? 
```{r}
# athens airbnb data (DATA RETRIEVED)
```
http://insideairbnb.com/get-the-data.html

Wow, can't wait to visit greece. also can't wait to learn more spatial analysis skills with R, really awesome! 


