---
title:
author: "Fisher Ankney"
date: "3/11/2018"
output: html_document
---

### Greece

insert image of greek flag here

One of my favorite countries, visited twice, plan on going again whenever I get the chance. Perfect opportuntiy to test out new found spatial analysis skills in R on one of my favorite destinations. Understanding greece using technologies in R. Who said you need GIS? 

```{r, message=F}
library('tidyverse')    
library('ggmap')
library('sp')
library('tmap')
library('raster')
library('rgdal')
library('tigris')
library('RColorBrewer')
library('viridisLite')
```

<br> 

If you forgot where Greece is
```{r, message=F}
greece <- c(lon = 21.8243, lat = 39.0742)

greece_far <- get_map(greece, zoom = 5, scale = 1, maptype="terrain")

ggmap(greece_far)
```

<br> 

Greece is broken into 11 administrative regions 

library(sp)
library(rgdal)

# Use dir() to find directory name
dir()

# Call dir() with directory name
dir("nynta_16c")

# Read in shapefile with readOGR(): neighborhoods
neighborhoods <- readOGR("nynta_16c", "nynta")

# summary() of neighborhoods
summary(neighborhoods)

```{r}
dir("input/GRC_adm")

grc_ad0 <- readOGR("input/GRC_adm", "GRC_adm0")
grc_ad1 <- readOGR("input/GRC_adm", "GRC_adm1")
grc_ad2 <- readOGR("input/GRC_adm", "GRC_adm2")
grc_ad3 <- readOGR("input/GRC_adm", "GRC_adm3")

summary(grc_ad0)
summary(grc_ad1)
summary(grc_ad2)
summary(grc_ad3)

tm_shape(grc_adm) + 
  tm_borders()
```
http://www.diva-gis.org/gdata

<br> 

But most people live in athens or thesoliniki
```{r}
# population distribution (DATA RETRIEVED)
```
http://www.diva-gis.org/gdata

<br> 

Greece is mostly mountainous, arid, amazing climate 
```{r}
# land cover / climate via raster? (DATA RETRIEVED)
```
http://www.diva-gis.org/gdata

<br> 

Perhaps The next time I visit I'll go to the biggest island of all, Crete. 
```{r}
crete <- c(lon = 24.8093, lat = 35.2401)

crete_far <- get_map(crete, zoom = 5, scale = 1, maptype="terrain")

ggmap(crete_far)

# Get map at zoom level 13: corvallis_map
crete_map <- get_map(crete, zoom = 8, scale = 1)

# Plot map at zoom level 13
ggmap(crete_map)

```


many archeology sites on crete
```{r}
# arceology in Crete (DATA RETRIEVED)
```
<br> 

and I better hurry because they're threatened by earthquakes
```{r}
# 
earthquakes and archeology sites 
```
https://www.apdkritis.gov.gr/el/dataset/%CE%B1%CF%81%CF%87%CE%B1%CE%B9%CE%BF%CE%B9-%CE%BD%CE%B1%CE%BF%CE%B9
https://www.apdkritis.gov.gr/el/dataset/%CE%B1%CF%81%CF%87%CE%B1%CE%B9%CE%BF%CE%BB%CE%BF%CE%B3%CE%B9%CE%BA%CE%BF%CE%B9-%CF%87%CF%89%CF%81%CE%BF%CE%B9
https://www.apdkritis.gov.gr/el/dataset/%CE%B1%CE%BD%CE%B1%CE%BA%CF%84%CE%BF%CF%81%CE%B1
earthquakes - https://www.apdkritis.gov.gr/el/dataset/%CF%83%CE%B5%CE%B9%CF%83%CE%BC%CE%BF%CE%B9-%CF%83%CF%84%CE%B7%CE%BD-%CE%BA%CF%81%CE%B7%CF%84%CE%B7

<br> 

but where to stay? I heard it hard to get a room in athens and it's really expensive, is this true? 
```{r}
# athens airbnb data (DATA RETRIEVED)
```
http://insideairbnb.com/get-the-data.html

Wow, can't wait to visit greece. also can't wait to learn more spatial analysis skills with R, really awesome! 


